---
title: "Visualizing the Napoleon's march"
author: "Partha Koundinya Panguluri"

toc: true
toc-title: On this page
code-fold: true
code-summary: "Show the code"
format:
  html:
    theme: united
    title-block-banner: true
    backgroundcolor: "#fcfcfc"
    embed-resources: true
---


Following resources were referred to understand Napoleon's march in ggplot2.\
[Minard meets ggplot2](http://euclid.psych.yorku.ca/www/psy6135/tutorials/Minard.html)\
[Exploring Minard's 1812 plot with ggplot2](https://www.andrewheiss.com/blog/2017/08/10/exploring-minards-1812-plot-with-ggplot2/)\
\
\
```{r}
#| echo: false
options(repos = c(CRAN = "http://cran.rstudio.com"))
```


# Part I - The Locations

## Setting off - Plot based on the geographical locations

```{r, echo=TRUE, message = FALSE, results='hide', warning=FALSE, fig.width=10,fig.height=3.5}

# Installing the libraries necessary

if(!require(pacman))
    install.packages("pacman")
pacman::p_load(gridExtra, scales, ggplot2, grid, dplyr)


# Importing the data
napoleon <- readRDS("data/napoleon.rds")
str(napoleon$troops)
str(napoleon$cities)
str(napoleon$temperatures)



#plotting a graph for napoleon_troops dataset with long and lat at x,y axes
ggplot(napoleon$troops, aes(long, lat)) +
  geom_path(aes(line_width = survivors))
```

The initial Napolean's March plot was difficult to understand. So, it was plotted with colors.\
Further modifications were made to improve the readability.\
Then, parameters like places, temperature were included to provide more insights about the event.

## Embedding aesthetics

```{r, echo=TRUE, message = FALSE, results='hide', warning=FALSE, fig.width=10,fig.height=3.5}

#  Plotting previous graph in different colours based on the path of troops
ggplot(napoleon$troops, aes(long, lat)) +
  geom_path(aes(size = survivors, colour = direction, group = group)) +
  coord_fixed()  #fixing coordinates based on aspect ratio
```

## Adjusting the scales

```{r, echo=TRUE, message = FALSE, results='hide', warning=FALSE, fig.width=10,fig.height=3.5}
#adjusting the scales to fit data into the graph
breaks <- c(1, 2, 3) * 10^5  #fixing the x-axis readability
ggplot(napoleon$troops, aes(long, lat)) +
  geom_path(aes(size = survivors, colour = direction, group = group), lineend = "round") +
  scale_size("Survivors", range = c(1,10), breaks = breaks, labels = scales::comma(breaks))+
  scale_color_hue("Direction", labels = c("Advance", "Retreat"))

#storing our plotted graph for troops
graph_troops <- last_plot()

```

## Plotting city locations

```{r, echo=TRUE, message = FALSE, results='hide', warning=FALSE, fig.width=10,fig.height=3.5}
#adding the cities along path of troops
graph_troops + geom_text(data = napoleon$cities, aes(label = city), size = 4)

#adding the cities to be adjusted into the graph as points
graph_troops + geom_point(data = napoleon$cities) +
  geom_text(data = napoleon$cities, aes(label = city), vjust = 1.5, family = "Times New Roman" )
```

## Avoiding any overlaps among points and city labels

```{r, echo=TRUE, message = FALSE, results='hide', warning=FALSE, fig.width=10,fig.height=3.5}
#using the ggrepel function to plot overlapping points/ path separately
install.packages("ggrepel")
library(ggplot2)
library(ggrepel)
graph_troops + geom_point(data = napoleon$cities) +
  geom_text_repel(data = napoleon$cities, aes(label = city))

#saving the plot with data from troops and cities dataset from the napoleon dataset 
graph_troops_cities <- last_plot()

```

# Part II - Temperature plots

## Presenting basic temperature plot

```{r, echo=TRUE, message = FALSE, results='hide', warning=FALSE, fig.width=10,fig.height=3.5}
#reading the temperature dataset to understand the data
napoleon$temperatures

#plotting line plot for temperature
ggplot(napoleon$temperatures, aes(long, temp)) +
  geom_path(color = "black", size = 1.5) +
  geom_point(size = 2)
```

## Inserting values

```{r, echo=TRUE, message = FALSE, results='hide', warning=FALSE, fig.width=10,fig.height=3.5}
#creating a label to show the temperature at different points on the graph
library(dplyr)
napoleon_temp <- napoleon$temperatures %>%
  mutate(format_degrees = paste0(temp, "° ", date))
head(napoleon_temp$format_degrees)

#plotting the graph with this label and format for temperature
ggplot(napoleon_temp, aes(long, temp)) +
  geom_path(color = "grey", size = 1.5) +
  geom_point(size = 1) +
  geom_text_repel(aes(label = format_degrees), size=2.5)

#saving the graph for temperature across different dates
graph_temp <- last_plot()

```

# Part I + Part II - The Assembly

```{r, echo=TRUE, message = FALSE, results='hide', warning=FALSE, fig.width=10,fig.height=3.5}
#compiling all the graphs together
library(gridExtra)
grid.arrange(graph_troops_cities, graph_temp)

#removing the labels for plotted figure for movement of troops and cities to match it with our temperature plot
library(ggplot2)
graph_troops_cities + coord_cartesian(xlim = c(24, 38)) +
  labs(x = NULL, y = NULL) + guides(color = FALSE, size = FALSE) +
  theme_void()

graph_troops_cities_edited <- last_plot()

#performing similar edits on the temperature graph and adjusting it
graph_temp + 
  coord_cartesian(xlim = c(24, 38)) +
  labs(x = NULL, y = "Temperature") +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(), axis.text.x = element_blank(), axis.ticks = element_blank(),
        panel.border = element_blank())

graph_temp_edited <- last_plot()

#compiling our edited graphs together
library(grid)
library(gridExtra)
grid.arrange(graph_troops_cities_edited, graph_temp_edited, nrow=2, heights=c(3.5, 1.2))
grid.rect(width = .99, height = .99, gp = gpar(lwd = 2, col = "gray", fill = NA))

```
